### 2016-10-20 docs/dg_penalty/main.tex
updating note

\pdfoutput=1

%\documentclass[preprint,10pt]{elsarticle}
\documentclass[10pt]{article}
%\documentclass[review]{siamart0216}

\usepackage{fullpage}
\usepackage{hyperref}
%\let\proof\relax
%\let\endproof\relax
\usepackage{amsmath,amssymb,amsfonts,mathrsfs}%,amsthm}
\usepackage[titletoc,toc,title]{appendix}

%\usepackage{lineno}
\usepackage{array} 
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

%% ====================================== alg package
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}
%% ====================================== end alg package

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{width=10cm,compat=1.9}
\pgfplotsset{
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}
\usepackage{stmaryrd}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\bs}[1]{\boldsymbol{#1}}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\tablab}[1]{\label{tab:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}

\newcommand{\theolab}[1]{\label{theo:#1}}
\newcommand{\theoref}[1]{\ref{theo:#1}}
\newcommand{\eqnlab}[1]{\label{eq:#1}}
\newcommand{\eqnref}[1]{\eqref{eq:#1}}
\newcommand{\seclab}[1]{\label{sec:#1}}
\newcommand{\secref}[1]{\ref{sec:#1}}
\newcommand{\lemlab}[1]{\label{lem:#1}}
\newcommand{\lemref}[1]{\ref{lem:#1}}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\Bh}{\widehat{B}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\Lk}{L^2\LRp{D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}

%\newenvironment{definition}[1][Definition]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
%\newenvironment{example}[1][Example]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vectwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vecthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

\newcommand{\note}[1]{{\color{blue}#1}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\tri}{{\rm tri}}
\newcommand{\sqr}{{\rm quad}}

\newcommand{\refhex}{\widehat{\mathcal{H}}}
\newcommand{\reftet}{\widehat{\mathcal{T}}}
\newcommand{\refwedg}{\widehat{\mathcal{W}}}

\newcommand{\refpyr}{\widehat{\mathcal{P}}}
\newcommand{\refpyrf}{\widehat{\mathcal{P}}^f}
\newcommand{\refpyrfq}{\widehat{\mathcal{P}}^{\sqr}}
\newcommand{\refpyrft}{\widehat{\mathcal{P}}^{\tri}}

\newcommand{\hex}{{\mathcal{H}}}
\newcommand{\tet}{{\mathcal{T}}}
\newcommand{\wedg}{{\mathcal{W}}}

\newcommand{\pyr}{{\mathcal{P}}}
\newcommand{\pyrf}{{\mathcal{P}}^f}
\newcommand{\pyrfq}{{\mathcal{P}}^{\sqr}}
\newcommand{\pyrft}{{\mathcal{P}}^{\tri}}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\half}{1/2}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\newcommand{\squareface}{[-1,1]^2}

\date{}
\author{Jesse Chan, T. Warburton}
\title{On penalty flux parameters in first order DG formulations}

\begin{document}

\maketitle
\tableofcontents

%\begin{abstract}
%
%\end{abstract}

\section{Introduction}

\section{DG numerical  fluxes}

We consider a first order system of hyperbolic equations
\[
\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\bm{F}_i(\bm{U})}{\bm{x}_i} = 0,
\]
which may alternatively be written as
\[
\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\LRp{\bm{A}_{i}U}}{\bm{x}_i}, \qquad \bm{A}_{i} = \pd{\bm{F}_i(\bm{U})}{\bm{U}}
\]
where $\bm{A}_{i}$ are symmetric matrices.  The semi-discrete DG formulation for such systems may be written as 
\[
\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{F}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{\bm{F}_n^*(\bm{U})\bm{n}_i,\bm{V}}_{\partial D^k}}
\]
where $\bm{n}$ is the outward normal on a face $f$ of $D^k$, and $\bm{F^*}_n$ is a numerical flux depending defined on shared faces between two elements.

%For convergence, $\bm{F^*}_n = \bm{F^*}_n(\bm{U})$ must be consistent such that, for exact solutions $\bm{U}$, 
%\[
%\bm{F^*}_n = \bm{F}(\bm{U})\cdot \bm{n}.
%\]

Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$, and let $\bm{F}^+, \bm{F}^-$ be evaluations of $\bm{F}(\bm{U})$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.  The average and jump of a vector valued function over $f$ are then defined as
\[
\avg{\bm{F}} = \frac{\bm{F}^+ + \bm{F}^-}{2}, \qquad \jump{\bm{F}} = \bm{F}^+ - \bm{F}^-.
\]
Typical numerical fluxes are defined as the sum of the average of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term
\begin{equation}
\bm{F^*}_n = \avg{\bm{F}(\bm{U})}\cdot \bm{n} - \bm{G}\LRp{\jump{\bm{U}}},
\label{eq:flux}
\end{equation}
where $\bm{G}$ is some convex function.

The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).  For some normal vector $\bm{n}$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}$.  By \note{add citation}, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
\[
\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^{-1}, \qquad \bm{\Lambda} = 
\left(\begin{array}{ccc}
\lambda_1 & & \\
& \ddots & \\
& & \lambda_d
\end{array}\right).
\]
For problems with continuous coefficients, the upwind numerical flux can be defined as
\[
\bm{F^*}_n = \bm{A}_n^+\bm{U}^- + \bm{A_n}^- \bm{U}^+,
\]
where the matrices $\bm{A_n}^+,\bm{A_n}^-$ are constructed from the positive and negative eigenvalues 
\begin{align*}
\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^{-1}\\
\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^{-1},
\end{align*}
and $\LRb{\bm{\Lambda}}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.  This can be rewritten as the sum of the central flux and a penalization
\[
\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1} \jump{\bm{U}}.  
\]
For problems with discontinuous coefficients, this may be generalized by solving a Riemann problem to yield
\[
\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1}\bm{U}}}.  
\]

An alternative to upwind fluxes are penalty fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
\[
\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{A}_n^T \bm{W} \jump{\bm{A}_n\bm{U}}.  
\]
where $\bm{W}$ is a diagonal weighting matrix with positive entries.  We take $\bm{W} = \tau \bm{I}$ in the following sections; however, the entries can be made to vary.  For example, taking $\bm{W} = \LRb{\bm{\Lambda}}^{-1}$ recovers the upwind fluxes \note{prove this}.
%\[
%\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{A}_n^T \LRb{\bm{Lambda}}^{-1} \jump{\bm{A}_n\bm{U}}.  
%\]

\subsection{Examples: acoustic wave equation}

Consider the isotropic acoustic wave equation in pressure-velocity form
\begin{align*}
\pd{p}{t} &= \Div \bm{u}\\
\pd{\bm{u}}{t} &= \Grad p.
\end{align*}
Let $\bm{U}$ denote the group variable $\bm{U} = (p,u,v)$, where $u$ and $v$ are the $x$ and $y$ components of velocity.  Then, in two dimensions, the isotropic wave equation is given as
\[
\pd{\bm{U}}{t} + \pd{\bm{A}_x\bm{U}}{x} + \pd{\bm{A}_y\bm{U}}{y} = 0, \qquad \bm{A}_x = 
\left(\begin{array}{ccc}
0 & 1 & 0\\
1 & 0 & 0\\
0 & 0 & 0
\end{array}
\right), \qquad 
\bm{A}_y = 
\left(\begin{array}{ccc}
0 & 0 & 1\\
0 & 0 & 0\\
1 & 0 & 0
\end{array}
\right).
\]
The normal flux matrix $\bm{A}_n$ is then
\[
\bm{A}_n = 
\left(\begin{array}{ccc}
0 & \bm{n}_x & \bm{n}_y\\
\bm{n}_x & 0 & 0\\
\bm{n}_y & 0 & 0
\end{array}
\right)
\]
implying that the penalty fluxes (with $\bm{W} = \tau\bm{I}$) are 
\[
\avg{\bm{A}_n \bm{U}} - \tau\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
\avg{\bm{u}_n}\\
\avg{p }\bm{n}_x\\
\avg{p }\bm{n}_y
\end{array}
\right) - 
\tau\left(\begin{array}{c}
\jump{p}\\
\jump{\bm{u_n}}\bm{n}_x\\
\jump{\bm{u_n}}\bm{n}_y
\end{array}
\right)
\]
For $\tau = 1$, these fluxes coincide with the upwind fluxes for isotropic media.  We note that the penalty flux affects only the normal component of the velocity; this is in contrast to Lax-Friedrichs-type fluxes, which penalize component-wise discontinuities in the velocity.

\note{Skew-symmetric discretization for constant coefficient problems \cite{kopriva2014energy}.  For wave propagation, can wrap variable coefficients into $A_0$ \cite{chan2016weight1,chan2016weight2}}.

\section{Dependence of spectra on the penalty parameter}

\note{Define spaces and show that $\bm{S}$ is symmetric definite based on DG discretization.}

We adapt the proof presented in \cite{Warburton20063205} to show that the spectra of $\bm{K}$ splits into two sets as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which moves left of the imaginary axis at a rate of $O(\tau)$.

We begin by splitting the DG approximation space into a conforming part $V^C$ and a non-conforming part $V^{NC}$.  Let $V$ be the DG approximation space
\[
V = \LRc{ u \in L^2(\Oh) : \left.u\right|_{D^k} \in P^N(D^k)}.  
\]
The conforming space $V^{C}$ is then defined as
\begin{equation}
V^{C} = \LRc{ u(\bm{x}) \in V: }
\label{eq:conf}
\end{equation}
We can similarly define the non-conforming approximation space $V^{NC}$ as the orthogonal complement of $V^{C}$, such that the full approximation space $V$ is 
\[
V = V^C \oplus V^{NC}.  
\]

This induces a block decomposition
\[
\bm{K} = \left(\begin{array}{cc}
\bm{A} & \bm{B}\\
-\bm{B}^T & \bm{C} + \tau \bm{S}
\end{array}\right).
\]
We first consider the spectra of $\bm{C} + \tau\bm{S}$.  By the construction of the DG formulation $\bm{C}$ is skew-symmetric, while $\bm{S}$ is symmetric and negative-definite.  Let $\lambda, \bm{u}$ be an eigenpair of $\bm{C} + \tau{\bm{S}}$ corresponding to the largest $N^D$ eigenvalues. Let $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.  Then, expanding and grouping terms in
\[
(\bm{C}+\tau\bm{S})(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
\]
we have that
\begin{align*}
(\bm{C}+\tau\bm{S})\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
(\bm{C}+\tau\bm{S})\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
\end{align*}
Assuming that $\bm{u}^*\bm{u} = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
%\begin{align*}
%\bm{v}^T\bm{A}\bm{v} &= \alpha\nor{\bm{v}}^2-\beta\bm{v}^T\bm{w}\\
%\bm{w}^T\bm{A}\bm{w} &= \beta\bm{w}^T{\bm{v}}+\alpha\nor{\bm{w}}^2.  
%\end{align*}
%\begin{align*}
%\bm{v}^T\bm{A}\bm{v} + \bm{w}^T\bm{A}\bm{w} &= \alpha\\
%\bm{v}^T\bm{A}\bm{w} - \bm{w}^T\bm{A}\bm{w} &= \beta.
%\end{align*}
using the skew-symmetry of $\bm{C}$ and symmetry of $\bm{S}$ yields
\begin{align*}
\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w} &= \beta\\
\tau\LRp{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} &= \alpha.
\end{align*}
Since $\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w}= 2\bm{w}^T\bm{C}\bm{v} = \beta$ is independent of $\tau$ and $\bm{v},\bm{w}$ are normalized, $\beta$ remains bounded as $\tau\rightarrow \infty$.  Similarly, since $\bm{S}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
\[
\LRb{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} = \LRb{\frac{\alpha}{\tau}}
\]
is bounded independently of $\tau$, and we conclude that $\LRb{\alpha} = O(\tau)$.  Thus, as $\tau$ increases, the eigenvalues of $\bm{C} + \tau\bm{S}$ are shifted further and further left of the imaginary axis.

We may now show how the spectra of $\bm{K}$ behaves as $\tau \rightarrow \infty$.  Following \cite{Warburton20063205}, let $\bm{U}$ and $\bm{Q}$ be unitary matrices whose columns contain the eigenvectors of $\bm{A}$ and $\bm{C} + \tau \bm{S}$, respectively.  Then, applying a block diagonal similarity transform yields
\[
\left(\begin{array}{cc}
\bm{U}^* & \\
& \bm{Q}^*
\end{array}\right)
\left(\begin{array}{cc}
\bm{A} & \bm{B}\\
-\bm{B}^T & \bm{C} + \tau \bm{S}
\end{array}\right)
\left(\begin{array}{cc}
\bm{U} & \\
& \bm{Q}
\end{array}\right)
 = \left(\begin{array}{cc}
\bm{\Lambda}_C & \bm{U}^T\bm{B}\bm{Q}\\
-\bm{Q}^T\bm{B}^T\bm{U} & \bm{\Lambda}_D
\end{array}\right),
\]
where $\bm{\Lambda}_C,\bm{\Lambda}_D$ are diagonal matrices whose entries consist of eigenvalues of $\bm{A},\bm{C} + \tau\bm{S}$, respectively.  Since $\bm{B}$ is independent of $\tau$, $\nor{\bm{U}^T\bm{B}\bm{Q}}$ can be bounded independently of $\tau$, assuming that $\bm{U}$ and $\bm{Q}$ are normalized.  Gerschgorin's theorem applied to $\bm{\Lambda}_C,\bm{\Lambda}_D$ then implies that the eigenvalues of $\bm{K}$ are contained in two sets of discs with radii independent of $\tau$.  The first set of discs are centered around $\lambda^C_i$ for $i = 1,\ldots,N^C$, while the second set of discs are centered around $\lambda^D_i$ for $i = 1,\ldots,N^D$, where ${\rm Re}(\lambda^D_i) = O(\tau)$.

Consider now the $N^C$ eigenvalues with the smallest magnitude.  By the Gerschgorin argument, these must remain bounded as $\tau \rightarrow \infty$.  Let $\bm{W} = (\bm{W}_C,\bm{W}_D)$ and $\bm{\Lambda}_C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
\[
\left(\begin{array}{cc}
\bm{A} & \bm{B}\\
-\bm{B}^T & \bm{C} + \tau \bm{S}
\end{array}\right)
\left(\begin{array}{c}
\bm{W}_C
\\
\bm{W}_D
\end{array}\right) = 
\left(\begin{array}{c}
\bm{A}\bm{W}_C + \bm{B}\bm{W}_D\\
-\bm{B}^T\bm{W}_C + \bm{C}\bm{W}_D + \tau \bm{S}\bm{W}_D
\end{array}\right)
= 
\bm{\Lambda}_C
\left(\begin{array}{c}
\bm{W}_C
\\
\bm{W}_D
\end{array}\right) 
\]
This implies 
\[
\tau \nor{\bm{S}\bm{W}_D} = \nor{\bm{\Lambda}_C \bm{W}_D + \bm{B}^T\bm{W}_C  - \bm{C}\bm{W}_D}.
\]
Since $\Lambda_C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}_C,\bm{W}_D$, implying that the non-conforming component $\bm{W}^D$ satisfies
\[
C\nor{\bm{W}_D} \leq \nor{\bm{S}\bm{W}_D} = O(1/\tau).
\]
As a consequence, $\bm{W}_D \rightarrow 0$ as $\tau\rightarrow \infty$, and the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the purely imaginary eigenvalues of $\bm{A}$ at a rate of $O(1/\tau)$.  These correspond to a discretization using the conforming approximation space (\ref{eq:conf}).

\section{Numerical experiments}

The above analysis illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.  However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau = O(1)$.  We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.

\subsection{1D experiments}

We consider first the scalar advection equation in 1D with periodic boundary conditions.  For $\tau = 0$, the eigenvalues of the DG discretization matrix are purely imaginary.  Figure~\ref{fig:track1D} shows the paths taken by these eigenvalues as $\tau$ increases from zero to $\tau = 4$.  As predicted, a subset of eigenvalues diverges towards the left half plane, which we refer to as \emph{divergent}.  The corresponding eigenmodes are shown in Figure~\ref{fig:trackmodes1D}, with inter-element jumps of these eigenfunctions increasing as $\tau$ increases.

\begin{figure}
\centering
\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.46\textwidth]{figs/trackedEigs.png}}
\hspace{1em}
\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.44\textwidth]{figs/trackedEigsZoom.png}}
\caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.  Eigenvalues are overlaid on these paths for $\tau = 0$, $\tau = 1$, and $\tau = 4$.  The zoomed in view near the imaginary axis shows the return of spurious modes to the imaginary axis for $\tau$ sufficiently large. }
\label{fig:track1D}
\end{figure}

\begin{figure}
\centering
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge1.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge2.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge3.png}}\\
\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge2.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge3.png}}
\caption{Behavior of divergent eigenvalues for advection as $\tau$ increases.  The order of approximation is $N=3$ on a mesh of 4 elements.} %The displayed eigenfunctions correspond to the highlighted eigenvalue.  }
\label{fig:trackmodes1D}
\end{figure}

However, for sufficiently large $\tau$, a subset of eigenvalues return to the imaginary axis.  Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a continuous function.  These eigenvalues which return to the imaginary axis correspond to a second set of \emph{spurious} modes, consisting of higher frequency components with sharp peaks, which have previously been observed in high order $C^0$ finite element methods \cite{ainsworth2014dispersive,hughes2014finite}.

\begin{figure}
\centering
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs1.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs2.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs3.png}}\\
\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModes1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModes2.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModes3.png}}
\caption{Return behavior of DG advection eigenvalues for sufficiently large $\tau$.  For $\tau = 1$ certain eigenvalues have negative real part, and the corresponding eigenmodes are damped.  For $\tau \gg 1$, certain eigenvalues return to the imaginary axis as spurious modes in conforming discretizations. }
\label{fig:spurious}
\end{figure}

We repeat the above experiment for the acoustic wave equation in 1D with reflection boundary conditions on $[-1,1]$, and compare against the exact eigenvalues $\lambda_i = i \frac{\pi}{2}$ for $i = 1,2,\ldots$.  The behavior of the eigenvalues as $\tau$ increases (Figure~\ref{fig:trackModesWave}) is similar to the behavior observed for the scalar advection equation.  An examination of the eigenmodes corresponding to divergent and spurious eigenvalues as $\tau \rightarrow \infty$ shows that these modes behave similarly to the spurious modes observed for the scalar advection equation, with individual pressure and velocity components converging to continuous functions.

%\begin{figure}
%\centering
%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs1.png}}
%\hspace{.5em}
%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs2.png}}
%\caption{ }
%\label{fig:trackModesWave}
%\end{figure}

\subsection{2D experiments}

In two space dimensions, taking $\tau\rightarrow \infty$ again results in spurious modes which return to the imaginary axis.  The corresponding eigenvalues converge to conforming functions.  For the acoustic wave equation, $p$ is continuous across faces, edges, and vertices, implying that $p\in H^1(\Oh)$, while only the normal component of $\bm{u}$ is continuous across faces, implying that $\bm{u} \in H({\rm div}; \Oh)$.

Uniform triangular mesh resulting from bisecting each element of a $2\times 2$ uniform quadrilateral mesh.  
%For the advection equation, 
%\begin{figure}
%\centering
%\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
%\hspace{.5em}
%\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
%\caption{Behavior of eigenvalues for advection }
%\label{fig:trackModesAdvec}
%\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{figs/waveEigs1.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{figs/waveEigs2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{figs/waveEigs3.png}}
\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions. Note the changing scale of the real axis.}
\label{fig:waveeigs}
\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}}
\hspace{2em}
\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}}
\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.  Divergent eigenvalues in the far left half plane are not shown.  }
\label{fig:trackEigsWave}
\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p2.png}}
\hspace{.5em}
\subfloat[$\tau = 50$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
\caption{Behavior of spurious returning modes for the acoustic wave equation in two dimensions.  }
\label{fig:trackModesWave}
\end{figure}

\section{Conclusions}

\bibliographystyle{unsrt}
\bibliography{dgpenalty}

\end{document}