### 2018-10-19 docs/skewform/deriv_estimate.tex

\pdfoutput=1

\documentclass{article}

\usepackage{cite}

\usepackage{amsmath, amssymb, amsfonts}
%\newtheorem*{remark}{Remark}
%\theoremstyle{assumption}
%\newtheorem{assumption}{Assumption}
\usepackage{hyperref}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{tikz}
\usepackage{hhline}
\usepackage{graphicx, color, bm}
\usepackage{subfig}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{slopetri}

\renewcommand{\hat}{\widehat}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\note}[1]{{\color{blue}{#1}}}

\graphicspath{{./figs/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\date{}
\author{}
\title{Error estimates for derivatives}

\begin{document}

Let $\LRp{\cdot,\cdot}, \LRp{\cdot,\cdot}_h$ denote the exact and quadrature-based $L^2$ inner product on the volume, and let $\LRa{\cdot,\cdot}, \LRa{\cdot,\cdot}_h$ denote the exact and quadrature-based $L^2$ inner products on the surface.  We will assume that, in $d > 1$ dimensions, the volume and surface quadratures satisfy
\begin{align*}
\LRb{\LRp{u,v} - \LRp{u,v}_h} \lesssim h^{r_{\rm vol}+1+d}  \nor{u}_{H^{r+1}\LRp{D^k}}\nor{v}_{H^{r+1}\LRp{D^k}}\\
\LRb{\LRa{u,v} - \LRa{u,v}_h} \lesssim h^{r_{\rm surf}+d} \nor{u}_{H^{r+1}\LRp{\partial D^k}}\nor{v}_{H^{r+1}\LRp{\partial D^k}}.
\end{align*}

We wish to show error estimates for certain finite difference approximations of derivatives.  More precisely, we will show that, for sufficiently regular $f$, the weak derivative $D_hf$ satisfies
\[
\nor{ \pd{f}{x} - {D}_h {f}}\leq Ch^{p+1} \nor{f}_{H^{p+1}}.
\]
More precisely, we will show that the $L^2$ projection satisfies
\[
\nor{ \Pi_N\pd{f}{x} - {D}_h {f}}\leq Ch^{\alpha} \nor{f}_{H^{p+1}}, \qquad \alpha = \min\LRp{r_{\rm vol} + 1 -p, r_{\rm surf} - p}.
\]

Main notes: if $\bm{Q}_N$ does not satisfy the SBP property, then the error estimate is limited by the accuracy of the weak derivative.  The strong derivative is always $O(h^p)$, but the weak derivative is $O\LRp{h^{\left\lfloor\alpha/2\right\rfloor}}$, where $\alpha$ is defined above.

\end{document}