### 2020-02-27 talks/WCCM2020/wccm2020-abstract.tex

\documentclass[a4paper]{wccm2020-abstract}
\usepackage{parskip}
\usepackage{pslatex}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\setlength{\parskip}{5pt}

\title{Mortar-based high order entropy stable discontinuous Galerkin schemes on non-conforming quadrilateral and hexahedral meshes}

\author{Jesse Chan$^{1}$}

\address{$^{1}$ Rice University (USA), 6100 Main St, Houston, TX, 77005, jesse.chan@rice.edu, jlchan.github.io}

\keywords{High order, discontinuous Galerkin, entropy stability} % Write down at least 3 Keywords

\begin{document}
\thispagestyle{empty}

High order discontinuous Galerkin (DG) methods provide improved accuracy and low numerical dispersion/dissipation for simulations of nonlinear conservation laws.  However, these methods also tend to suffer from instability in practice, requiring filtering, limiting, or artificial dissipation to prevent solution blow up.  \textit{Entropy stable} nodal DG methods based on summation-by-parts (SBP) operators and flux differencing address this instability by ensuring satisfaction of a semi-discrete entropy inequality.  Entropy stable DG methods typically utilize Lobatto quadrature points; however, recent work has extended such methods to more general nodal points \cite{esdg1,esdg3}.  In this talk, we will present entropy stable DG methods based on Gauss points, which can improve accuracy on warped and non-conforming meshes \cite{esdg2}.

\begin{thebibliography}{99}
\bibitem{esdg1} Chan, Jesse. "On discretely entropy conservative and entropy stable discontinuous Galerkin methods." Journal of Computational Physics 362 (2018): 346-374.
\bibitem{esdg2}  Chan, Jesse, David C. Del Rey Fern\'{a}ndez, and Mark H. Carpenter. ``Efficient entropy stable Gauss collocation methods.'' SIAM Journal on Scientific Computing 41.5 (2019): A2938-A2966.
\bibitem{esdg3} Chan, J. Skew-Symmetric Entropy Stable Modal Discontinuous Galerkin Formulations. Journal of Scientific Computing 81, 459–485 (2019) doi:10.1007/s10915-019-01026-w
\end{thebibliography}

\end{document}