### 2020-04-03 docs/ESDG_ROM/letter.tex

\documentclass{letter}
\usepackage{fullpage}
\signature{Jesse Chan}
\address{6100 Main St,\\
Houston, TX 77005}

\begin{document}
\begin{letter}{Dr Remi Abgrall\\JCP Editor}

\opening{Dear Dr Abgrall,}

Please find enclosed the manuscript, 
\begin{center}
Jesse Chan\\
\textit{Entropy stable reduced order modeling of nonlinear conservation laws},
\end{center}
which we would like to submit for publication as a research paper in JCP.

The submitted manuscript describes the construction projection-based reduced order models which are globally conservative and satisfy a discrete conservation or dissipation of entropy.  The reduced models are constructed by combining appropriate hyper-reduction strategies with techniques from entropy stable finite volume and summation by parts (SBP) schemes.  The novelty of the proposed approach is in the discretization of the convective nonlinearity, which adapts the ``flux differencing'' technique from SBP schemes to reduced order models.

We hope that the method and results discussed in this manuscript would appeal to the readership of JCP.  The author confirms that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}

\end{document}